### 2023-08-01 paper/paper.tex
paper

\documentclass[12pt, letterpaper, twocolumn]{article}
\title{Title}
\author{Author}
\date{Date}

\begin{document}
\maketitle

\section{Introduction}

\section{Related Work}

\section{Methods}

\onecolumn
\begin{table}[h]
\centering
\begin{tabular}{|c c c|} 
    \hline
    Name & input length & embedding dimension  \\ 
    \hline
    embedding reproduction (rep)        & 64    & 64    \\
    dnabert                             & 512     & 768  \\
    prose                               & 545   & 100    \\
    alphafold                           & 1024   & 18744 \\
    evolutionary scale modeling (esm)   & 1024    & 1280 \\  
    \hline
\end{tabular}
\caption{Table to test captions and labels.}
\label{table:data}
\end{table}
\twocolumn

\section{Results}

\end{document}

### 2023-08-02 paper/paper.tex
papaer

@@
 \documentclass[12pt, letterpaper, twocolumn]{article}
+\usepackage[utf8]{inputenc}
+\usepackage{hyperref}
+\usepackage{graphicx}
+\usepackage{placeins}
+
 \title{Title}
@@
 \centering
-\begin{tabular}{|c c c|} 
+\begin{tabular}{|l c c|} 
     \hline
@@
     \hline
-    embedding reproduction (rep)        & 64    & 64    \\
-    dnabert                             & 512     & 768  \\
-    prose                               & 545   & 100    \\
-    alphafold                           & 1024   & 18744 \\
-    evolutionary scale modeling (esm)   & 1024    & 1280 \\  
+    embedding reproduction (rep)\cite{yang2018learned}       & 64    & 64    \\
+    dnabert \cite{ji2021dnabert}                     & 512     & 768  \\
+    prose   \cite{bepler2021learning}                   & 512   & 100    \\
+    alphafold  \cite{jumper2021highly}                 & 1024   & 384 \\
+    evolutionary scale modeling (esm2) \cite{lin2022language}   & 1024    & 1280 \\  
     \hline
 \end{tabular}
-\caption{Table to test captions and labels.}
+\caption{Embedders used in the experiments.}
 \label{table:data}
@@
 
+\onecolumn
+\newpage
+\FloatBarrier
+\bibliographystyle{plain}
+\bibliography{bibliography}
+
 \end{document}

### 2023-08-02 pipeline_enrichment.py
tests

@@ -43,10 +43,10 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
     et.add_multistage(
         function=pipeline_build_embeddings_matrix,
         list_args=[   
-            {"embedder" : "rep", "combiner_method" : "pca" },
-            {"embedder" : "rep", "combiner_method" : "average" },
-            {"embedder" : "rep", "combiner_method" : "sum" },
-            {"embedder" : "rep", "combiner_method" : "max" },
+            # {"embedder" : "rep", "combiner_method" : "pca" },
+            # {"embedder" : "rep", "combiner_method" : "average" },
+            # {"embedder" : "rep", "combiner_method" : "sum" },
+            # {"embedder" : "rep", "combiner_method" : "max" },
             
             {"embedder" : "dnabert", "combiner_method" : "pca" },
             {"embedder" : "dnabert", "combiner_method" : "average" },
@@ -135,15 +135,15 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         function=pipeline_build_embeddings_linkage_matrix,
         list_args=[
         {"metric" : "euclidean", "method" : "average"},
-        {"metric" : "euclidean", "method" : "complete"},
-        {"metric" : "euclidean", "method" : "ward"},
-        {"metric" : "euclidean", "method" : "centroid"},
-        {"metric" : "euclidean", "method" : "single"},
-        {"metric" : "euclidean", "method" : "median"},
+        # {"metric" : "euclidean", "method" : "complete"},
+        # {"metric" : "euclidean", "method" : "ward"},
+        # {"metric" : "euclidean", "method" : "centroid"},
+        # {"metric" : "euclidean", "method" : "single"},
+        # {"metric" : "euclidean", "method" : "median"},
         
-        {"metric" : "cosine", "method" : "average"},
-        {"metric" : "cosine", "method" : "complete"},
-        {"metric" : "cosine", "method" : "single"},
+        # {"metric" : "cosine", "method" : "average"},
+        # {"metric" : "cosine", "method" : "complete"},
+        # {"metric" : "cosine", "method" : "single"},
         ]
     )
 
@@ -161,6 +161,8 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         annotation_dict = {}
         
         for record in records:
+            # TODO this line cause problems because the prefix of the name is not always "sp" and is not possible to 
[…]

### 2023-08-02 pipeline_phylogenetic.py

@@ -65,7 +65,7 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
             {"embedder" : "esm", "combiner_method" : "average" },
             {"embedder" : "esm", "combiner_method" : "sum" },
             {"embedder" : "esm", "combiner_method" : "max" },
-        
+            
         ]
     )
 
@@ -101,8 +101,9 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
     et.add_multistage(
         function=pipeline_pca,
         list_args=[
-            {"n_components": "default"},
             {"n_components": "all"},
+            {"n_components": "default"},
+
         ]
     )
 
@@ -220,7 +221,7 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         for i in range(predict_labels_matrix.shape[1]):
             adjusted_rand_scores.append(adjusted_rand_score(predict_labels_matrix[:,i], gtrue_labels_matrix[:,i]))
         
-        return {"mean_adjusted_rand_score" : np.mean(adjusted_rand_scores), "max_adjusted_rand_score" : np.max(adjusted_rand_scores)}
+        return {"mean_adjusted_rand_score" : np.mean(adjusted_rand_scores)}
         
 
 
@@ -236,11 +237,11 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
 
 if __name__ == "__main__":
     
-    # EMBEDDINGS_PATH = "./dataset/globins/globins.json"
-    # GROUND_TRUE_PATH = "./dataset/globins/globins.dist"
-
-    EMBEDDINGS_PATH =  "./dataset/emoglobina/embeddings"
+    EMBEDDINGS_PATH = "./dataset/emoglobina/embeddings"
     GROUND_TRUE_PATH = "./dataset/emoglobina/emoglobina.dist"
+
+    # EMBEDDINGS_PATH =  "./dataset/batterio/embeddings"
+    # GROUND_TRUE_PATH = "./dataset/batterio/batterio.dist"
     
     
     et = main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH)

### 2023-08-15 paper/paper.tex
paper

@@
 
+In this section we describe the pipeline used to analyze the embeddings.
+As shown in Table \ref{table:data}, the input length is different between the models as well as the output produced.
+We want to address the following problems: 1) compare different methods to join togheter the amminoacid-specific contextual representations in order to have a representation for the whole chunk; 2) compare different methods to join togheter the representations of the chunks in order to have a representation for the whole protein; 3) find out if these representations reflect known properties of the proteins.
+
+\subsection{Single amminoacid contextual representation}
+
 \onecolumn
@@
     \hline
-    Name & input length & embedding dimension  \\ 
+    Name & input length (chunk) & embedding dimension  \\ 
     \hline
-    embedding reproduction (rep)\cite{yang2018learned}       & 64    & 64    \\
-    dnabert \cite{ji2021dnabert}                     & 512     & 768  \\
-    prose   \cite{bepler2021learning}                   & 512   & 100    \\
-    alphafold  \cite{jumper2021highly}                 & 1024   & 384 \\
-    evolutionary scale modeling (esm2) \cite{lin2022language}   & 1024    & 1280 \\  
+    embedding reproduction (rep)\cite{yang2018learned}       & 64    & 64 per chunk   \\
+    dnabert \cite{ji2021dnabert}                     & 512     & 768 per chunk \\
+    prose   \cite{bepler2021learning}                   & 512   & 100 per amino acid   \\
+    alphafold  \cite{jumper2021highly}                 & 1024   & 384 per ammino acid\\
+    evolutionary scale modeling (esm2) \cite{lin2022language}   & 1024    & 1280 per ammino acid \\  
     \hline

### 2023-09-04 autoembedding/combiners.py
solved todo

@@ -2,7 +2,6 @@ import numpy as np
 from sklearn.decomposition import PCA
 from sklearn.preprocessing import StandardScaler
 
-# TODO rewrite also the embedders to be compatible with the combiners
 
 def combiner(raw_embedding : np.array , method: str)-> np.array :
     """

### 2023-09-04 autoembedding/results_manager.py

@@ -25,7 +25,6 @@ def results2table(
                 computations_dict[combiner][pca][embedder] = None
     
     # fill the dict
-    # TODO add the weght of the enrichment method in the preferred method
     for result, pipeline in r:  # for each result and the pipeline that generated it
         combiner = None
         pca = None

### 2023-09-04 autoembedding/utils.py

@@ -49,20 +49,7 @@ def newick_to_linkage(newick: str):
     Convert newick tree into scipy linkage matrix
 
     :param newick: newick file path or string
-    :param label_order: list of labels, e.g. ['A', 'B', 'C']
     :returns: linkage matrix
-    """
-
-    # TODO check why the new implementation dosn't read the newick file generated by FastTree
-    # TODO why the old implementation works with unrooter trees and the new one doesn't
-
-    """ new implementation 
-    pdm = PhyloDM.load_from_newick_path(newick)
-
-    dm = pdm.dm(norm=False)
-    labels = pdm.taxa()
-
-    return linkage(squareform(dm)), labels
     """
 
     # newick string -> cophenetic_matrix

### 2023-09-04 build_phylotree.py

@@ -1,13 +1,0 @@
-
-# TODO make a proper script
-# TODO remeber that 0 is fully equals, the max value is completely different
-
-
-# ./clustalo-1.2.4-Ubuntu-x86_64 -i dataset/globins/globins.fasta -o dataset/globins/globins.aln2 --distmat-out=dataset/globis/globins.dist --full --force
-
-
-# protein by default, nucleotide with -nt
-
-# ./FastTree dataset/emoglobina/emoglobina.aln > dataset/emoglobina/emoglobina.dnd
-
-# TODO add the rooting of the tree if needed (depends on the implementation that reads the tree, the new one seems to need the rooting)

### 2023-09-04 pipeline_enrichment.py

@@ -11,7 +11,6 @@ import numpy as np
 from Bio import SeqIO
 from autoembedding.results_manager import results2file
 
-# TODO check if the execution tree works well swapping the order of hyperparamenters
 
 def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):    
     

### 2023-09-05 paper/paper.tex
paper

@@
 \subsection{Single amminoacid contextual representation}
+We tried four methods to join togheter the amminoacid embeddings in order to produce a fixed size embedding for the chunk: average, maximum, sum and principal component analysis (PCA).
+Note that even if these operators are commutative, the overall process do takes into account the order of the amminoacids precisely because the embeddings are contextual.
+
+The same operator used to combine the amminoacid embeddings is also used to combine the embeddings of the chunks of the sequence.
 
@@
 \end{tabular}
-\caption{Embedders used in the experiments.}
+\caption{Embedders used in the experiments, their maximum imput length and the dimension of the embedding produced.}
 \label{table:data}

### 2023-09-06 paper/paper.tex
paper

@@
 
-\subsection{Single amminoacid contextual representation}
+\subsection{Combining the contextual representations}
 We tried four methods to join togheter the amminoacid embeddings in order to produce a fixed size embedding for the chunk: average, maximum, sum and principal component analysis (PCA).
@@
 The same operator used to combine the amminoacid embeddings is also used to combine the embeddings of the chunks of the sequence.
+
+\subsection{Comparison with known informations}
+Given a set of embeddings of sequences we want to analyze their distribution in the embedding space comparing it with both the distance matrix produced during the multiple sequence alignment with Clustal Omega \cite{sievers2011fast} and protein annotations as gene ontology, UniProt keywords and taxonimy \cite{uniprot23}.
+
+\subsubsection{Allignment distance}
+In order to compare the distances between the sequences in the embedding space with the alignment distance we performed an agglomerative clustering on both the matrices, the resulting tree is then cut at each levels: flat partitions of all possibles number of clusters are produced.
+We perform a comparison of the partitions with the same number of clusters using the adjusted rand score \cite{hubert1985comparing}.
+The mean of these score, starting from two clusters up to $ \#elements - 1 $ clusters is called mean adjusted rand score (MARS).
 

### 2023-09-08 paper/paper.tex
paper

--- 2023-09-05 paper/paper.tex
@@
 In this section we describe the pipeline used to analyze the embeddings.
-As shown in Table \ref{table:data}, the input length is different between the models as well as the output produced.
+As shown in Table \ref{tab:embedders}, the input length is different between the models as well as the output produced.
 We want to address the following problems: 1) compare different methods to join togheter the amminoacid-specific contextual representations in order to have a representation for the whole chunk; 2) compare different methods to join togheter the representations of the chunks in order to have a representation for the whole protein; 3) find out if these representations reflect known properties of the proteins.
 
-\subsection{Single amminoacid contextual representation}
+\subsection{Combining the contextual representations}
 We tried four methods to join togheter the amminoacid embeddings in order to produce a fixed size embedding for the chunk: average, maximum, sum and principal component analysis (PCA).
@@
 
+\subsection{Comparison with known informations}
+Given a set of embeddings of sequences we want to analyze their distribution in the embedding space comparing it with both the distance matrix produced during the multiple sequence alignment with Clustal Omega \cite{sievers2011fast} and annotations as Gene Ontology \cite{10.1093/genetics/iyad031, ashburner2000gene}, UniProtKB Keywords and NCBI Taxonimy \cite{uniprot23}.
+
+\subsubsection{Allignment distance \label{sec:allignment}}
+In order to compare the distances between the sequences in the embedding space with the alignment distance we performed an agglomerative clustering on both the matrices, the resulting tree is then cut at each level: flat partitions of all possibles number of clusters are produced.
+We performed a comparison of the partitions with the same number of clusters using the adjusted rand score \cite{hubert1985comparing}.
+The mean of these score, starting from two clusters up to $ \#elements - 1 $ clusters is called mean adjusted rand score (MARS).
+
+\subsubsection{Enrichment analysis}
+We wanted to analyze the properties of the embeddings also at an higher level.
+The Gene Ontology (GO) describes our knowledge of the sequence with respect to: molecular function, cellular component and biological process; there ara also more specific controlled vocabulary as the UniProt Keywords and hierarchical classifications specific for sequences as the NCBI Taxonomy.
+
+Whatever they are the sets of words to describe the sequences in our datasets, we want to build a distance between sequences among them.
+Given $A$ and $B$ the sets of annotations of two sequences we computed the distance in two possible ways: $$d1 = \frac{2 * |A \cap B|}{|A| + |B|} $$ $$d2 = \max\{ \frac{|A \cap B|}{|A|}, \frac{|A \cap B|}{|B|} \} $$
+
+Both of them vary between $0$ and $1$, however $d1$ goes to 1 only when the two sets are equals while $d2$ goes to 1 also when one set is a subset of the other.
+After calculating one of these distance between all possible pair in the dataset we end up with a similarity matrix, that can be easly transformed in a distance matrix that is possible to comprare with the distance matrix derived from the distance between the embeddings using the MARS as described in subsection \ref{sec:allignment}.
+
 \onecolumn
-\begin{table}[h]
+\begin{table}[ht]
+\centering
+\begin{tabular}{|p{3cm} p{3cm} c c c |} 
+    \hline
+    Name & description & number of sequences & type & avg length   \\ 
+    \hline
+    hemoglobin &  hemoglobin for various organisms & 761 & amminoacid & 142  \\
+    \hline
+    mouse & mouse proteome     & 974 & amminoacid & 516 \\
+    \hline
+    bacterium & bacterium proteome  & 259 &  amminoacid & 427  \\
+    \hline
+    covid19 & covid19 complete genome & 77 & nucleotides & 29831  \\
+    \hline
+    meningitis & meningitis complete genome & 68 & nucleotides & 2240049 \\
+    \hline
+\end{tabular}
+\caption{Datasets used in the experiments.}
+\label{tab:dataset}
+\end{table}
+\twocolumn
+
+\onecolumn
+\begin{table}[ht]
 \centering
@@
 \caption{Embedders used in the experiments, their maximum imput length and the dimension of the embedding produced.}
-\label{table:data}
+\label{tab:embedders}
 \end{table}

### 2023-09-20 paper/paper.tex
paper

--- 2023-09-05 paper/paper.tex
@@
 \section{Introduction}
-
-\section{Related Work}
+In the latest few years, machine learning and deep learning have revolutionized the natural language processing field (NLP)\cite{khurana2023natural}.
+At the same time some fundamental ideas developed in NLP have been succesfuly applied to another type of language, the biological one: DNA, RNA and amminoacid sequences bringing to excellent results even in the complex task of protein structure prediction\cite{jumper2021highly, lin2022language}.
+One of these fundamental ideas is word embedding \cite{mikolov2013efficient} because it transforms words into points in space, therefore easy to process.
+It is well known that sequence similarity do not always correspond to functional similarity \cite{kosloff2008sequence}.
 
@@
 In this section we describe the pipeline used to analyze the embeddings.
-As shown in Table \ref{table:data}, the input length is different between the models as well as the output produced.
-We want to address the following problems: 1) compare different methods to join togheter the amminoacid-specific contextual representations in order to have a representation for the whole chunk; 2) compare different methods to join togheter the representations of the chunks in order to have a representation for the whole protein; 3) find out if these representations reflect known properties of the proteins.
+As shown in Table \ref{tab:embedders}, the input length is different between the models as well as the output produced.
+We want to address the following problems: 1) compare different methods to join togheter the amminoacid-specific contextual representations in order to have a representation for the whole chunk and subsequently join togheter the representations of the chunks in order to have a representation for the whole protein; 2) find out if these representations reflect known properties of the proteins.
 
-\subsection{Single amminoacid contextual representation}
+\subsection{Represent protein sequences as continuous vectors}
+short description of the main characteristics of each embedder 
+\subsubsection{SeqVec}
+
+\subsection{Combining the (contextual) representations}
 We tried four methods to join togheter the amminoacid embeddings in order to produce a fixed size embedding for the chunk: average, maximum, sum and principal component analysis (PCA).
-Note that even if these operators are commutative, the overall process do takes into account the order of the amminoacids precisely because the embeddings are contextual.
 
@@
 
+Note on the contextal embedding: seqvec: contextual
+
+\subsection{Comparison with known informations}
+Given a set of embeddings of sequences we want to analyze their distribution in the embedding space comparing it with both the distance matrix produced during the multiple sequence alignment with Clustal Omega \cite{sievers2011fast} and higher level annotations as Gene Ontology \cite{10.1093/genetics/iyad031, ashburner2000gene}, UniProtKB Keywords and NCBI Taxonimy \cite{uniprot23}.
+
+\subsubsection{Similarity between distance matrices \label{sec:similarity}}
+In order to compare two distance matrices we performed an agglomerative clustering on both, the resulting tree is then cut at each level obtaining flat partitions of all possibles number of clusters.
+We performed a pairwise comparison of the partitions having the same number of clusters using the adjusted rand score \cite{hubert1985comparing}.
+The mean of these score, starting from two clusters up to $ \#elements - 1 $ clusters is called mean adjusted rand score (MARS).
+We compared different distance metrics as well as different methods to perform the hierarchical clustering.
+
+\subsubsection{Enrichment analysis}
+The alignment distance matrix provide an evolutionary related distance between sequences \cite{SOFI202247}, we also wanted to analyze the properties of the embeddings at an higher level.
+The Gene Ontology (GO) describes our knowledge of the sequence with respect to: molecular function, cellular component and biological process; there ara also more specific controlled vocabulary as the UniProt Keywords and hierarchical classifications specific for sequences as the NCBI Taxonomy.
+
+Whatever they are the sets of words to describe the sequences in our datasets, we want to build a distance between sequences among them.
+Given $A$ and $B$ the sets of annotations of two sequences we computed the distance in two possible ways: $$d1 = \frac{2 * |A \cap B|}{|A| + |B|} $$ $$d2 = \max\{ \frac{|A \cap B|}{|A|}, \frac{|A \cap B|}{|B|} \} $$
+
+Both of them vary between $0$ and $1$, however $d1$ goes to 1 only when the two sets are equals while $d2$ goes to 1 also when one set is a subset of the other.
+After calculating one of these distance between all possible pair in the dataset we end up with a similarity matrix, that can be easly transformed in a distance matrix that is possible to comprare with the distance matrix derived from the distance between the embeddings using the MARS as described in subsection \ref{sec:similarity}.
+
 \onecolumn
-\begin{table}[h]
+\begin{table}[htb]
 \centering
@@
 \caption{Embedders used in the experiments, their maximum imput length and the dimension of the embedding produced.}
-\label{table:data}
+\label{tab:embedders}
+\end{table}
+\twocolumn
+
+\onecolumn
+\begin{table}[htb]
+\centering
+\begin{tabular}{|p{3cm} p{3cm} c c c |} 
+    \hline
+    Name & description & number of sequences & type & avg length   \\ 
+    \hline
+    hemoglobin &  hemoglobin for various organisms & 761 & amminoacids & 142  \\
+    \hline
+    mouse & mouse proteome     & 974 & amminoacids & 516 \\
+    \hline
+    bacterium & bacterium proteome  & 259 &  amminoacids & 427  \\
+    \hline
+    covid19 & covid19 complete genome & 77 & nucleotides & 29831  \\
+    \hline
+    meningitis & meningitis complete genome & 68 & nucleotides & 2240049 \\
+    \hline
+\end{tabular}
+\caption{Datasets used in the experiments.}
+\label{tab:dataset}
 \end{table}
@@
 \section{Results}
+\subsection{Phylogenetic}
+\subsection{Enrichment}
+\subsection{Projections?}
+\subsection{Classification?}
+\subsection{Pointwise representation similarity?}
 

### 2023-09-20 pipeline_phylogenetic.py
seqvec test

@@ -45,6 +45,11 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
             # {"embedder" : "rep", "combiner_method" : "average" },
             # {"embedder" : "rep", "combiner_method" : "sum" },
             # {"embedder" : "rep", "combiner_method" : "max" },
+
+            {"embedder" : "seqvec", "combiner_method" : "pca" },
+            {"embedder" : "seqvec", "combiner_method" : "average" },
+            {"embedder" : "seqvec", "combiner_method" : "sum" },
+            {"embedder" : "seqvec", "combiner_method" : "max" },
             
             {"embedder" : "dnabert", "combiner_method" : "pca" },
             {"embedder" : "dnabert", "combiner_method" : "average" },

### 2023-09-23 paper/paper.tex
paper

@@
 short description of the main characteristics of each embedder
+
+\subsubsection{prose}
+learning with strictural supervision
+
+\subsubsection{alphafold}
+
+- is a deep learning system not a language model
+
+- predict residue-residue distances from sequence families and fold proteins based on the predicted distance constraints
+
+- rely on large datasets of protein sequences that are similar enough to be aligned with high confidence but contain enough divergence to confidently infer statistical couplings between positions
+
+- unable to learn patterns across large-scale databases of possibly unrelated proteins and have limited ability to draw on the increasing structure and function information available
+
 \subsubsection{SeqVec}
+To build their model Heinzinger et al.\cite{heinzinger2019modeling} adapted the standar ELMo configuration \cite{peters-etal-2018-deep} to work with protein sequences modifying the number of tokens and the unroll steps.
+It is composed by 1 CharCNN and 2 LSTM-Layers.
+Given a protein sequence of arbitrary length it returns 3072 features for each residue derived by concatenating the outputs of the three layers of ELMo, each describing a token with a vector of length 1024.
+In order to obtain a smaller representation for each amminoacid i computed the mean of the three layers (as also suggested in the official repository).
+Given the architecture of the ELMo, these representation are contextual-dependent.
 
-\subsection{Combining the (contextual) representations}
+\subsection{Combining the (contextual) representations}derived by concatenating the outputs of the three layers of ELMo, each describing a token with a vector of length 1024
 We tried four methods to join togheter the amminoacid embeddings in order to produce a fixed size embedding for the chunk: average, maximum, sum and principal component analysis (PCA).
@@
     embedding reproduction (rep)\cite{yang2018learned}       & 64    & 64 per chunk   \\
+    seqvec \cite{heinzinger2019modeling} & 1024 & 1024 per amminoacid \\
     dnabert \cite{ji2021dnabert}                     & 512     & 768 per chunk \\

### 2023-09-23 autoembedding/combiners.py

@@ -2,6 +2,8 @@ import numpy as np
 from sklearn.decomposition import PCA
 from sklearn.preprocessing import StandardScaler
 
+# TODO implement t-sne
+
 
 def combiner(raw_embedding : np.array , method: str)-> np.array :
     """

### 2023-09-23 pipeline_phylogenetic.py

@@ -99,7 +99,7 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         return { "embeddings_matrix" : embeddings_matrix, "embeddings_IDs": embeddings_IDs}
 
     et.add_multistage(
-        function=pipeline_pca,
+        function=pipeline_pca, # TODO t-sne
         list_args=[
             {"n_components": "all"},
             {"n_components": "default"},

### 2023-09-29 paper/paper.tex
paper

@@
 It is well known that sequence similarity do not always correspond to functional similarity \cite{kosloff2008sequence}.
+Application of protein language model embeddings to downstream tasks was first demostrated by Bepler and Berger (2019) \cite{bepler2019learning}
 
@@
 
-\subsubsection{prose}
-learning with strictural supervision
+\subsubsection{prose \cite{bepler2021learning}}
+The model structure is a multi-layer bidirectional Long Short Term Memory (bi-LSTM).
+Following the intuition that some aspect of proteins structure and semantic can never be discoverable by statistical sequence models alone, they came up with the idea of multi-task learning with strictural supervision.
+The learning tasks are: 1) classical masked language modeling task, 2) residue-residue contact prediction, 3) structural similarity prediction.
+This model ouperformed existing approaches in both transmembrane position labeling and phenotypes prediction of sequence variants.
 
@@
 Given a protein sequence of arbitrary length it returns 3072 features for each residue derived by concatenating the outputs of the three layers of ELMo, each describing a token with a vector of length 1024.
-In order to obtain a smaller representation for each amminoacid i computed the mean of the three layers (as also suggested in the official repository).
+In order to obtain a smaller representation for each amminoacid we computed the mean of the three layers (as also suggested in the official repository).
 Given the architecture of the ELMo, these representation are contextual-dependent.
+
+\subsubsection{Evolutionary scale modeling}
 

### 2023-10-06 paper/paper.tex
paper

@@
 
-\subsubsection{prose \cite{bepler2021learning}}
+\subsubsection{Prose \cite{bepler2021learning}}
 The model structure is a multi-layer bidirectional Long Short Term Memory (bi-LSTM).
@@
 
-\subsubsection{alphafold}
+\subsubsection{Alphafold}
+
+- evoformer module and structure module. The Evoformer, builds separate MSA and residue-pairwise embedding spaces
 
@@
 
+- used without the MSA and templates
+
 \subsubsection{SeqVec}
@@
 
-\subsubsection{Evolutionary scale modeling}
+\subsubsection{Evolutionary scale modeling \cite{lin2023evolutionary}}
+The underlying architecture is a BERT \cite{devlin2018bert} style encoder tarnsformer with modifications in the number of layers, number of attention heads, hidden size and feed forward hidden size.
+A further improvement is the use of Rotary Position Embedding \cite{su2021roformer} that allows to generalize beyond the context window it is trained on.
+The main advantages compared to AlphaFold are the removal of the the need of multiple sequence alignment and an increasing up to two order of magnitude of  the speed of the prediction pipeline.
+We used the version with 33 layers and 650B parameters (compare with alphafold)
 
-\subsection{Combining the (contextual) representations}derived by concatenating the outputs of the three layers of ELMo, each describing a token with a vector of length 1024
+\subsection{Combining the (contextual) representations}
 We tried four methods to join togheter the amminoacid embeddings in order to produce a fixed size embedding for the chunk: average, maximum, sum and principal component analysis (PCA).
@@
     alphafold  \cite{jumper2021highly}                 & 1024   & 384 per ammino acid\\
-    evolutionary scale modeling (esm2) \cite{lin2022language}   & 1024    & 1280 per ammino acid \\  
+    evolutionary scale modeling (esm2) \cite{lin2023evolutionary}   & 1024    & 1280 per ammino acid \\  
     \hline

### 2023-10-09 paper/paper.tex
paper

@@
 
-\subsubsection{Alphafold}
+\subsubsection{Alphafold \cite{jumper2021highly}}
+This is the only approach that do not relay on a language model but is a combination of a bioinformatics and physical approaches.
+It relys on large datasets of protein sequences that are similar enough to be aligned with high confidence but contain enough divergence to confidently infer statistical couplings between positions.
+It consists in two modules, Evoformer module and structure module.
+The Evoformer builds separate MSA and residue-pairwise embedding spaces.
+As it is structured, this system is not able to learn patterns across large-scale databases of possibly unrelated proteins \cite{bepler2021learning}.
+In the experiments we used the ``single'' representation of the sequence, without the MSA and the templates.
 
-- evoformer module and structure module. The Evoformer, builds separate MSA and residue-pairwise embedding spaces
+Leggi altro paper sul genoma
 
-- is a deep learning system not a language model
-
-- predict residue-residue distances from sequence families and fold proteins based on the predicted distance constraints
-
-- rely on large datasets of protein sequences that are similar enough to be aligned with high confidence but contain enough divergence to confidently infer statistical couplings between positions
-
-- unable to learn patterns across large-scale databases of possibly unrelated proteins and have limited ability to draw on the increasing structure and function information available
-
-- used without the MSA and templates
-
-\subsubsection{SeqVec}
+\subsubsection{SeqVec \cite{heinzinger2019modeling} }
 To build their model Heinzinger et al.\cite{heinzinger2019modeling} adapted the standar ELMo configuration \cite{peters-etal-2018-deep} to work with protein sequences modifying the number of tokens and the unroll steps.

### 2023-10-12 pipeline_enrichment.py
bugfix

@@ -18,7 +18,7 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
 
     # BUILD EMBEDDING MATRIX (WITH COMBINER)
     def pipeline_build_embeddings_matrix(previous_stage_output : dict, embedder: str, combiner_method : str) -> dict:
-        
+
         """
         Built the embeddings matrix from the embeddings dict
 
@@ -102,8 +102,12 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
     et.add_multistage(
         function=pipeline_pca,
         list_args=[
-            {"n_components": "default"},
-            {"n_components": "all"},
+            {"n_components": 10},
+            {"n_components": 20},
+            {"n_components": 30},
+            {"n_components": 40},
+            {"n_components": 50},
+            {"n_components": 'all'},
         ]
     )
 
@@ -134,15 +138,15 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         function=pipeline_build_embeddings_linkage_matrix,
         list_args=[
         {"metric" : "euclidean", "method" : "average"},
-        # {"metric" : "euclidean", "method" : "complete"},
-        # {"metric" : "euclidean", "method" : "ward"},
-        # {"metric" : "euclidean", "method" : "centroid"},
-        # {"metric" : "euclidean", "method" : "single"},
-        # {"metric" : "euclidean", "method" : "median"},
+        {"metric" : "euclidean", "method" : "complete"},
+        {"metric" : "euclidean", "method" : "ward"},
+        {"metric" : "euclidean", "method" : "centroid"},
+        {"metric" : "euclidean", "method" : "single"},
+        {"metric" : "euclidean", "method" : "median"},
         
-        # {"metric" : "cosine", "method" : "average"},
-        # {"metric" : "cosine", "method" : "complete"},
-        # {"metric" : "cosine", "method" : "single"},
+        {"metric" : "cosine", "method" : "average"},
+        {"metric" : "cosine", "method" : "complete"},
+        {"metric" : "cosine", "method" : "single"},
         ]
     )
 
@@ -160,18 +164,24 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         annotation_dict = {}
         
         for record in records:
-            # TODO this line cause problems because the prefix of the name is not always "sp" and is not possible to 
[…]

### 2023-10-12 pipeline_phylogenetic.py

@@ -135,15 +135,15 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         function=pipeline_build_embeddings_linkage_matrix,
         list_args=[
             {"metric" : "euclidean", "method" : "average"},
-            # {"metric" : "euclidean", "method" : "complete"},
-            # {"metric" : "euclidean", "method" : "ward"},
-            # {"metric" : "euclidean", "method" : "centroid"},
-            # {"metric" : "euclidean", "method" : "single"},
-            # {"metric" : "euclidean", "method" : "median"},
+            {"metric" : "euclidean", "method" : "complete"},
+            {"metric" : "euclidean", "method" : "ward"},
+            {"metric" : "euclidean", "method" : "centroid"},
+            {"metric" : "euclidean", "method" : "single"},
+            {"metric" : "euclidean", "method" : "median"},
             
-            # {"metric" : "cosine", "method" : "average"},
-            # {"metric" : "cosine", "method" : "complete"},
-            # {"metric" : "cosine", "method" : "single"},
+            {"metric" : "cosine", "method" : "average"},
+            {"metric" : "cosine", "method" : "complete"},
+            {"metric" : "cosine", "method" : "single"},
         ]
     )
 
@@ -170,15 +170,15 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         fixed_args={"ground_true_path" : GROUND_TRUE_PATH},
         list_args=[
             {"metric" : "euclidean", "method" : "average"},
-            # {"metric" : "euclidean", "method" : "complete"},
-            # {"metric" : "euclidean", "method" : "ward"},
-            # {"metric" : "euclidean", "method" : "centroid"},
-            # {"metric" : "euclidean", "method" : "single"},
-            # {"metric" : "euclidean", "method" : "median"},
+            {"metric" : "euclidean", "method" : "complete"},
+            {"metric" : "euclidean", "method" : "ward"},
+            {"metric" : "euclidean", "method" : "centroid"},
+            {"metric" : "euclidean", "method" : "single"},
+            {"metric" : "euclidean", "method" : "median"},
             
-            # {"metric" : "cosine", "method" : "average"},
[…]

### 2023-10-12 pipeline_enrichment.py
new experiments

@@ -151,7 +151,7 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
     )
 
     
-    def pipeline_build_gt_linkage_matrix(previous_stage_output : dict, metric, method, edge_weight, ground_true_path)-> dict:
+    def pipeline_build_gt_linkage_matrix(previous_stage_output : dict, metric, method, edge_weight, use_go, use_keywords, use_taxonomy, ground_true_path)-> dict:
 
         embeddings_linkage_matrix = previous_stage_output["embeddings_linkage_matrix"]
         embeddings_IDs = previous_stage_output["embeddings_IDs"]
@@ -184,20 +184,26 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
                 name_j = name_j[name_j.find("|")+1:]
                 
                 # the amount of common annotations between the two sequences, i.e. A inter B
-                capacity =\
-                    len(set(annotation_dict[name_i]['go']).intersection(set(annotation_dict[name_j]['go']))) +\
-                    len(set(annotation_dict[name_i]['keywords']).intersection(set(annotation_dict[name_j]['keywords'])))+\
-                    len(set(annotation_dict[name_i]['taxonomy']).intersection(set(annotation_dict[name_j]['taxonomy'])))
+                capacity = 0
+
                 # the amount of annotations of the first sequence
-                A =\
-                    len(set(annotation_dict[name_i]['go'])) +\
-                    len(set(annotation_dict[name_i]['keywords'])) +\
-                    len(set(annotation_dict[name_i]['taxonomy']))
+                A = 0
                 # the amount of annotations of the second sequence
-                B =\
-                    len(set(annotation_dict[name_j]['go'])) +\
-                    len(set(annotation_dict[name_j]['keywords'])) +\
-                    len(set(annotation_dict[name_j]['taxonomy']))
+                B = 0
+
+                if use_go:
+                    capacity += len(set(annotation_dict[name_i]['go']).intersection(set(annotation_dict[name_j]['go'])))
+                    A += len(set(annotation_dict[name_i]['go']))
+                    B += len(set(annotation_dict[name_j]['go'])) 
+                if use_keywords:
+                    capacity += len(set(annotation_dict[name_i]['keywords']).intersection(set(annotation_dict[name_j]['keywords'])))
+                    A += len(set(annotation_dict[name_i]['keywords']))
[…]

### 2023-10-12 pipeline_phylogenetic.py

@@ -106,9 +106,12 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
     et.add_multistage(
         function=pipeline_pca, # TODO t-sne
         list_args=[
-            {"n_components": "all"},
-            {"n_components": "default"},
-
+            {"n_components": 10},
+            {"n_components": 20},
+            {"n_components": 30},
+            {"n_components": 40},
+            {"n_components": 50},
+            {"n_components": "all"},       
         ]
     )
 
@@ -242,14 +245,14 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
 
 if __name__ == "__main__":
     
-    # EMBEDDINGS_PATH = "./dataset/emoglobina/embeddings"
-    # GROUND_TRUE_PATH = "./dataset/emoglobina/emoglobina.dist"
+    EMBEDDINGS_PATH = "./dataset/emoglobina/embeddings"
+    GROUND_TRUE_PATH = "./dataset/emoglobina/emoglobina.dist"
 
     # EMBEDDINGS_PATH =  "./dataset/batterio/embeddings"
     # GROUND_TRUE_PATH = "./dataset/batterio/batterio.dist"
 
-    EMBEDDINGS_PATH =  "./dataset/topo/embeddings"
-    GROUND_TRUE_PATH = "./dataset/topo/topo.dist"
+    # EMBEDDINGS_PATH =  "./dataset/topo/embeddings"
+    # GROUND_TRUE_PATH = "./dataset/topo/topo.dist"
     
     
     et = main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH)

### 2023-10-12 autoembedding/results_manager.py
paper

@@ -14,7 +14,7 @@ def results2table(
 
     combiners = ["pca", "average", "sum", "max"]
     pcas = ["all", "default"]
-    embedders = ["rep", "dnabert", "prose", "alphafold", "esm"	]
+    embedders = ["dnabert", "seqvec", "prose", "alphafold", "esm"	]
 
     # generate the empty dict with the argument that are relevant for the table
     for combiner in combiners:

### 2023-10-14 autoembedding/results_manager.py
new results

@@ -13,7 +13,7 @@ def results2table(
     computations_dict = {} # dict[combiner][pca][embedder] = score
 
     combiners = ["pca", "average", "sum", "max"]
-    pcas = ["all", "default"]
+    pcas = [ '10', '20', '30', '40', '50', 'all']
     embedders = ["dnabert", "seqvec", "prose", "alphafold", "esm"	]
 
     # generate the empty dict with the argument that are relevant for the table
@@ -35,7 +35,7 @@ def results2table(
                 embedder = args["embedder"]
                 combiner = args["combiner_method"]
             if stage == "pipeline_pca":
-                pca = args["n_components"]
+                pca = str(args["n_components"])
             if stage == "pipeline_build_embeddings_linkage_matrix":
                 metric_embedding = args["metric"]
                 method_embedding = args["method"]

### 2023-10-14 pipeline_phylogenetic.py

@@ -245,14 +245,14 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
 
 if __name__ == "__main__":
     
-    EMBEDDINGS_PATH = "./dataset/emoglobina/embeddings"
-    GROUND_TRUE_PATH = "./dataset/emoglobina/emoglobina.dist"
+    # EMBEDDINGS_PATH = "./dataset/emoglobina/embeddings"
+    # GROUND_TRUE_PATH = "./dataset/emoglobina/emoglobina.dist"
 
     # EMBEDDINGS_PATH =  "./dataset/batterio/embeddings"
     # GROUND_TRUE_PATH = "./dataset/batterio/batterio.dist"
 
-    # EMBEDDINGS_PATH =  "./dataset/topo/embeddings"
-    # GROUND_TRUE_PATH = "./dataset/topo/topo.dist"
+    EMBEDDINGS_PATH =  "./dataset/topo/embeddings"
+    GROUND_TRUE_PATH = "./dataset/topo/topo.dist"
     
     
     et = main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH)

### 2023-10-14 paper/paper.tex
paper

@@
 \usepackage{placeins}
+\usepackage{booktabs}
+\usepackage{multirow}
 
@@
 In the latest few years, machine learning and deep learning have revolutionized the natural language processing field (NLP)\cite{khurana2023natural}.
-At the same time some fundamental ideas developed in NLP have been succesfuly applied to another type of language, the biological one: DNA, RNA and amminoacid sequences bringing to excellent results even in the complex task of protein structure prediction\cite{jumper2021highly, lin2022language}.
+At the same time some fundamental ideas developed in NLP have been successfully applied to another type of language, the biological one: DNA, RNA and amminoacid sequences bringing to excellent results even in the complex task of protein structure prediction\cite{jumper2021highly, lin2022language}.
 One of these fundamental ideas is word embedding \cite{mikolov2013efficient} because it transforms words into points in space, therefore easy to process.
 It is well known that sequence similarity do not always correspond to functional similarity \cite{kosloff2008sequence}.
-Application of protein language model embeddings to downstream tasks was first demostrated by Bepler and Berger (2019) \cite{bepler2019learning}
+Application of protein language model embeddings to downstream tasks was first demonstrated by Bepler and Berger (2019) \cite{bepler2019learning}
 
@@
 
-In this section we describe the pipeline used to analyze the embeddings.
-As shown in Table \ref{tab:embedders}, the input length is different between the models as well as the output produced.
-We want to address the following problems: 1) compare different methods to join togheter the amminoacid-specific contextual representations in order to have a representation for the whole chunk and subsequently join togheter the representations of the chunks in order to have a representation for the whole protein; 2) find out if these representations reflect known properties of the proteins.
+In this section we describe the pipeline used to analyze the embeddings, i.e.
+vectors in a high-dimensional real space that represent a chunk of a genetic sequence (of a hole one if it is short enough).
+As shown in Table \ref{tab:embedders}, the input length accepted is different between different models as well as the dimensionality of the output produced.
+The maximum length of the input sequence is usually determined by the length of the examples the model was training with or by limitations due to the structure of the model such as the size of the attention module in the models that use it.
+These numerical representation of sequences are then used to various downstream applications.
+
+Because of their structure, these model output a vector for each single item of the input sequence (amminoacid or nucleotides) so we end up with a matrix representation of a sequence with different number of rows for each different sequence.
+Futhermore, if the sequence that we want to analyze is longer then the model's limit we have to split it in subsequences and thus the representation became a tensor.
+
+We want to address the following problems: 1) compare different methods to join together the amminoacid-specific contextual representations in order to have a representation for the whole chunk and subsequently join together the representations of the chunks in order to have a representation for the whole protein; 2) find out if these representations reflect known properties of the proteins.
+
+It is important to note that the vectorial representation of an item of the sequence (nucleotides or amminoacid) is not fixed, it is instead contextual, i.e.
+it depends, theoretically, on each other item in the sequence.
+This is intuitively good in natural language because a word can assume different meaning in different context \cite{salant2017contextualized}, and this is also good for nucleotides or amminoacids in biological sequences.
 
 \subsection{Represent protein sequences as continuous vectors}
-short description of the main characteristics of each embedder
+
+\subsubsection{DNABert}
 
@@
 The model structure is a multi-layer bidirectional Long Short Term Memory (bi-LSTM).
-Following the intuition that some aspect of proteins structure and semantic can never be discoverable by statistical sequence models alone, they came up with the idea of multi-task learning with strictural supervision.
+Following the intuition that some aspect of proteins structure and semantic can never be discoverable by statistical sequence models alone, they came up with the idea of multi-task learning with structural supervision.
 The learning tasks are: 1) classical masked language modeling task, 2) residue-residue contact prediction, 3) structural similarity prediction.
-This model ouperformed existing approaches in both transmembrane position labeling and phenotypes prediction of sequence variants.
+This model outperformed existing approaches in both transmembrane position labeling and phenotypes prediction of sequence variants.
 
@@
 This is the only approach that do not relay on a language model but is a combination of a bioinformatics and physical approaches.
-It relys on large datasets of protein sequences that are similar enough to be aligned with high confidence but contain enough divergence to confidently infer statistical couplings between positions.
+It rely on large datasets of protein sequences that are similar enough to be aligned with high confidence but contain enough divergence to confidently infer statistical couplings between positions.
 It consists in two modules, Evoformer module and structure module.
@@
 
-Leggi altro paper sul genoma
-
 \subsubsection{SeqVec \cite{heinzinger2019modeling} }
-To build their model Heinzinger et al.\cite{heinzinger2019modeling} adapted the standar ELMo configuration \cite{peters-etal-2018-deep} to work with protein sequences modifying the number of tokens and the unroll steps.
+To build their model Heinzinger et al.\cite{heinzinger2019modeling} adapted the standard ELMo configuration \cite{peters-etal-2018-deep} to work with protein sequences modifying the number of tokens and the unroll steps.
 It is composed by 1 CharCNN and 2 LSTM-Layers.
@@
 \subsubsection{Evolutionary scale modeling \cite{lin2023evolutionary}}
-The underlying architecture is a BERT \cite{devlin2018bert} style encoder tarnsformer with modifications in the number of layers, number of attention heads, hidden size and feed forward hidden size.
+The underlying architecture is a BERT \cite{devlin2018bert} style encoder transformer with modifications in the number of layers, number of attention heads, hidden size and feed forward hidden size.
 A further improvement is the use of Rotary Position Embedding \cite{su2021roformer} that allows to generalize beyond the context window it is trained on.
@@
 \subsection{Combining the (contextual) representations}
-We tried four methods to join togheter the amminoacid embeddings in order to produce a fixed size embedding for the chunk: average, maximum, sum and principal component analysis (PCA).
+We tried four methods to join together the amminoacid embeddings in order to produce a fixed size embedding for the chunk: average, maximum, sum and principal component analysis (PCA).
 
@@
 
-Note on the contextal embedding: seqvec: contextual
-
-\subsection{Comparison with known informations}
-Given a set of embeddings of sequences we want to analyze their distribution in the embedding space comparing it with both the distance matrix produced during the multiple sequence alignment with Clustal Omega \cite{sievers2011fast} and higher level annotations as Gene Ontology \cite{10.1093/genetics/iyad031, ashburner2000gene}, UniProtKB Keywords and NCBI Taxonimy \cite{uniprot23}.
+\subsection{Comparison with known information}
+Given a set of embeddings of sequences we want to analyze their distribution in the embedding space comparing it with both the distance matrix produced during the multiple sequence alignment with Clustal Omega \cite{sievers2011fast} and higher level annotations as Gene Ontology \cite{10.1093/genetics/iyad031, ashburner2000gene}, UniProtKB Keywords and NCBI Taxonomy \cite{uniprot23}.
 
@@
 We performed a pairwise comparison of the partitions having the same number of clusters using the adjusted rand score \cite{hubert1985comparing}.
-The mean of these score, starting from two clusters up to $ \#elements - 1 $ clusters is called mean adjusted rand score (MARS).
+The mean of these score, starting from two clusters up to $ \#elements - 1 $ clusters is called mean adjusted rand index (MARI).
 We compared different distance metrics as well as different methods to perform the hierarchical clustering.
+EXPLAIN BETTER
 
@@
 Both of them vary between $0$ and $1$, however $d1$ goes to 1 only when the two sets are equals while $d2$ goes to 1 also when one set is a subset of the other.
-After calculating one of these distance between all possible pair in the dataset we end up with a similarity matrix, that can be easly transformed in a distance matrix that is possible to comprare with the distance matrix derived from the distance between the embeddings using the MARS as described in subsection \ref{sec:similarity}.
+After calculating one of these distance between all possible pair in the dataset we end up with a similarity matrix, that can be easily transformed in a distance matrix that is possible to compare with the distance matrix derived from the distance between the embeddings using the MARS as described in subsection \ref{sec:similarity}.
 
-\onecolumn
-\begin{table}[htb]
+\begin{table*}[htb]
 \centering
@@
 \end{tabular}
-\caption{Embedders used in the experiments, their maximum imput length and the dimension of the embedding produced.}
+\caption{Embedders used in the experiments, their maximum input length and the dimension of the embedding produced.}
 \label{tab:embedders}
-\end{table}
-\twocolumn
+\end{table*}
 
-\onecolumn
-\begin{table}[htb]
+\begin{table*}[htb]
 \centering
@@
 \label{tab:dataset}
-\end{table}
-\twocolumn
+\end{table*}
 
@@
 \subsection{Phylogenetic}
+
+\begin{table*}[hbt]
+    \begin{tabular}{llrrrr}
+        \toprule
+         &  & seqvec & prose & alphafold & esm \\
+        combiner & dimensional PCA &  &  &  &  \\
+        \midrule
+        \multirow[t]{6}{*}{pca} & 10 & 0.449140 & 0.351432 & 0.417911 & 0.495245 \\
+         & 20 & 0.494195 & 0.395091 & 0.468677 & 0.545581 \\
+         & 30 & 0.509105 & 0.389776 & 0.478603 & 0.584235 \\
+         & 40 & 0.533046 & 0.388992 & 0.488803 & 0.589202 \\
+         & 50 & 0.544053 & 0.391228 & 0.489178 & 0.590380 \\
+         & all & 0.564945 & 0.394987 & 0.519607 & 0.605519 \\
+        \cline{1-6}
+        \multirow[t]{6}{*}{average} & 10 & 0.440643 & 0.376631 & 0.427137 & 0.511643 \\
+         & 20 & 0.493682 & 0.407182 & 0.467401 & 0.554667 \\
+         & 30 & 0.519014 & 0.407446 & 0.481296 & 0.578104 \\
+         & 40 & 0.534462 & 0.408016 & 0.488328 & 0.588670 \\
+         & 50 & 0.544056 & 0.407953 & 0.486588 & 0.601595 \\
+         & all & 0.577440 & 0.406862 & 0.351767 & 0.610861 \\
+        \cline{1-6}
+        \multirow[t]{6}{*}{sum} & 10 & 0.437402 & 0.382771 & 0.437270 & 0.493417 \\
+         & 20 & 0.480627 & 0.411288 & 0.474695 & 0.536754 \\
+         & 30 & 0.499964 & 0.407558 & 0.480758 & 0.568252 \\
+         & 40 & 0.525452 & 0.407691 & 0.490450 & 0.572879 \\
+         & 50 & 0.524109 & 0.409393 & 0.497554 & 0.583100 \\
+         & all & 0.558963 & 0.405949 & 0.359855 & 0.605848 \\
+        \cline{1-6}
+        \multirow[t]{6}{*}{max} & 10 & 0.429831 & 0.415769 & 0.374368 & 0.514790 \\
+         & 20 & 0.464959 & 0.498250 & 0.397569 & 0.579752 \\
+         & 30 & 0.477753 & 0.506698 & 0.415304 & 0.587261 \\
+         & 40 & 0.487472 & 0.514747 & 0.428992 & 0.595889 \\
+         & 50 & 0.499499 & 0.522068 & 0.435171 & 0.611635 \\
+         & all & 0.542751 & 0.533704 & 0.378067 & 0.680641 \\
+        \cline{1-6}
+        \bottomrule
+        \end{tabular}
+
+\caption{Emoglobin Phylogenetic results.}
+\label{tab:emoglobin_phylo_results}
+\end{table*}
+
+\begin{table*}[hbt]
+    \begin{tabular}{llrrrr}
+        \toprule
+         &  & seqvec & prose & alphafold & esm \\
+        combiner & dimensional PCA &  &  &  &  \\
+        \midrule
+        \multirow[t]{6}{*}{pca} & 10 & 0.190254 & 0.212593 & 0.123817 & 0.154119 \\
+         & 20 & 0.223326 & 0.215994 & 0.158653 & 0.211017 \\
+         & 30 & 0.233909 & 0.216776 & 0.169543 & 0.236152 \\
+         & 40 & 0.248802 & 0.216817 & 0.165458 & 0.259220 \\
+         & 50 & 0.255239 & 0.216956 & 0.170329 & 0.268735 \\
+         & all & 0.284410 & 0.212964 & 0.166672 & 0.284101 \\
+        \cline{1-6}
+        \multirow[t]{6}{*}{average} & 10 & 0.203596 & 0.234889 & 0.133038 & 0.194540 \\
+         & 20 & 0.252383 & 0.238390 & 0.160997 & 0.267515 \\
+         & 30 & 0.265740 & 0.238646 & 0.168218 & 0.294172 \\
+         & 40 & 0.282326 & 0.239766 & 0.170593 & 0.308133 \\
+         & 50 & 0.293596 & 0.239742 & 0.170439 & 0.318216 \\
+         & all & 0.337058 & 0.239850 & 0.126641 & 0.330147 \\
+        \cline{1-6}
+        \multirow[t]{6}{*}{sum} & 10 & 0.158431 & 0.215487 & 0.151187 & 0.141865 \\
+         & 20 & 0.168831 & 0.220564 & 0.154411 & 0.169305 \\
+         & 30 & 0.181248 & 0.221456 & 0.165493 & 0.186215 \\
+         & 40 & 0.182201 & 0.221607 & 0.162561 & 0.187705 \\
+         & 50 & 0.183775 & 0.222080 & 0.160401 & 0.190762 \\
+         & all & 0.195369 & 0.214632 & 0.112511 & 0.207289 \\
+        \cline{1-6}
+        \multirow[t]{6}{*}{max} & 10 & 0.213904 & 0.285951 & 0.088717 & 0.280855 \\
+         & 20 & 0.276499 & 0.307956 & 0.122844 & 0.342551 \\
+         & 30 & 0.294216 & 0.316039 & 0.135043 & 0.380750 \\
+         & 40 & 0.310357 & 0.316190 & 0.143899 & 0.397330 \\
+         & 50 & 0.319428 & 0.328722 & 0.154333 & 0.413625 \\
+         & all & 0.271817 & 0.334124 & 0.099899 & 0.438918 \\
+        \cline{1-6}
+        \bottomrule
+        \end{tabular}       
+\caption{Mouse Phylogenetic results.}
+\label{tab:mouse_phylo_results}
+\end{table*}
+
+\begin{table*}[hbt]
+    \begin{tabular}{llrrrr}
+        \toprule
+         &  & seqvec & prose & alphafold & esm \\
+        combiner & dimensional PCA &  &  &  &  \\
+        \midrule
+        \multirow[t]{6}{*}{pca} & 10 & 0.064957 & 0.072183 & 0.012768 & 0.034683 \\
+         & 20 & 0.085704 & 0.076620 & 0.005497 & 0.050977 \\
+         & 30 & 0.091166 & 0.076632 & 0.005371 & 0.052120 \\
+         & 40 & 0.090972 & 0.076629 & 0.008708 & 0.052657 \\
+         & 50 & 0.096276 & 0.076608 & 0.022336 & 0.058888 \\
+         & all & 0.092587 & 0.072544 & 0.007800 & 0.086005 \\
+        \cline{1-6}
+        \multirow[t]{6}{*}{average} & 10 & 0.063419 & 0.089908 & 0.017157 & 0.035883 \\
+         & 20 & 0.078590 & 0.089790 & 0.011811 & 0.053720 \\
+         & 30 & 0.087748 & 0.089808 & 0.007368 & 0.053789 \\
+         & 40 & 0.092050 & 0.089814 & 0.008073 & 0.054903 \\
+         & 50 & 0.094008 & 0.089814 & 0.006989 & 0.066082 \\
+         & all & 0.099638 & 0.092655 & 0.010589 & 0.084124 \\
+        \cline{1-6}
+        \multirow[t]{6}{*}{sum} & 10 & 0.017759 & 0.060247 & -0.000334 & 0.005171 \\
+         & 20 & 0.020039 & 0.059339 & 0.004637 & 0.007319 \\
+         & 30 & 0.022856 & 0.059442 & 0.005062 & 0.007984 \\
+         & 40 & 0.024805 & 0.059451 & 0.005057 & 0.008785 \\
+         & 50 & 0.024293 & 0.059436 & 0.004787 & 0.008381 \\
+         & all & 0.025664 & 0.060248 & 0.000751 & 0.023070 \\
+        \cline{1-6}
+        \multirow[t]{6}{*}{max} & 10 & 0.035823 & 0.074083 & 0.000905 & 0.034337 \\
+         & 20 & 0.042759 & 0.088803 & 0.003758 & 0.064972 \\
+         & 30 & 0.039637 & 0.091899 & 0.000352 & 0.081378 \\
+         & 40 & 0.038814 & 0.093861 & -0.000994 & 0.091238 \\
+         & 50 & 0.036684 & 0.098092 & -0.000269 & 0.093263 \\
+         & all & 0.031241 & 0.097953 & 0.010153 & 0.112645 \\
+        \cline{1-6}
+        \bottomrule
+        \end{tabular}        
+\caption{Bacterium Phylogenetic results.}
+\label{tab:bacterium_phylo_results}
+\end{table*}
+
 \subsection{Enrichment}

### 2023-10-14 autoembedding/results_manager.py

@@ -2,6 +2,9 @@ import pandas as pd
 
 def results2table(
     r, 
+    embedders = ["dnabert", "seqvec", "prose", "alphafold", "esm"],
+    combiners = ["pca", "average", "sum", "max"],
+    pcas = [ '10', '20', '30', '40', '50', 'all'],
     metric="mean_adjusted_rand_score", 
     preferred_metric_embedding="euclidean",
     preferred_method_embedding="ward",
@@ -11,10 +14,6 @@ def results2table(
     ):
     
     computations_dict = {} # dict[combiner][pca][embedder] = score
-
-    combiners = ["pca", "average", "sum", "max"]
-    pcas = [ '10', '20', '30', '40', '50', 'all']
-    embedders = ["dnabert", "seqvec", "prose", "alphafold", "esm"	]
 
     # generate the empty dict with the argument that are relevant for the table
     for combiner in combiners:

### 2023-10-15 autoembedding/results_manager.py
running experiments

@@ -11,6 +11,7 @@ def results2table(
     preferred_metric_gt="euclidean",
     preferred_method_gt="ward", 
     preferred_edge_weight="method_1",
+    preferred_annotation = "go"
     ):
     
     computations_dict = {} # dict[combiner][pca][embedder] = score
@@ -28,6 +29,8 @@ def results2table(
         combiner = None
         pca = None
         embedder = None  
+        edge_weight = preferred_edge_weight
+        annotation = preferred_annotation
         for stage, args in pipeline: # for each stage and the arguments that were passed to it
             # keep the arguments that are relevant for the table
             if stage == "pipeline_build_embeddings_matrix":
@@ -45,14 +48,25 @@ def results2table(
                 try:
                     edge_weight = args["edge_weight"]
                 except:
-                    edge_weight = preferred_edge_weight
+                    pass
+
+                try:
+                    if args['use_go']:
+                        annotation = "go"
+                    if args['use_keywords']:
+                        annotation = "keywords"
+                    if args['use_taxonomy']:
+                        annotation = "taxonomy"
+                except:
+                    pass
                     
         
         if  metric_embedding == preferred_metric_embedding and \
             method_embedding == preferred_method_embedding and \
             metric_gt == preferred_metric_gt and \
             method_gt == preferred_method_gt and \
-            edge_weight == preferred_edge_weight:
+            edge_weight == preferred_edge_weight and \
+            annotation == preferred_annotation:
         
             computations_dict[combiner][pca][embedder] = result[metric]
 

### 2023-10-15 pipeline_enrichment.py

@@ -46,6 +46,11 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
             # {"embedder" : "rep", "combiner_method" : "average" },
             # {"embedder" : "rep", "combiner_method" : "sum" },
             # {"embedder" : "rep", "combiner_method" : "max" },
+
+            {"embedder" : "seqvec", "combiner_method" : "pca" },
+            {"embedder" : "seqvec", "combiner_method" : "average" },
+            {"embedder" : "seqvec", "combiner_method" : "sum" },
+            {"embedder" : "seqvec", "combiner_method" : "max" },
             
             {"embedder" : "dnabert", "combiner_method" : "pca" },
             {"embedder" : "dnabert", "combiner_method" : "average" },
@@ -103,10 +108,10 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         function=pipeline_pca,
         list_args=[
             {"n_components": 10},
-            {"n_components": 20},
             {"n_components": 30},
-            {"n_components": 40},
             {"n_components": 50},
+            {"n_components": 70},
+            {"n_components": 90},
             {"n_components": 'all'},
         ]
     )
@@ -352,14 +357,14 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
 if __name__ == "__main__":
     
 
-    # EMBEDDINGS_PATH =  "./dataset/batterio/embeddings"
-    # GROUND_TRUE_PATH = "./dataset/batterio/batterio.xml"
+    EMBEDDINGS_PATH =  "./dataset/batterio/embeddings"
+    GROUND_TRUE_PATH = "./dataset/batterio/batterio.xml"
 
     # EMBEDDINGS_PATH =  "./dataset/emoglobina/embeddings"
     # GROUND_TRUE_PATH = "./dataset/emoglobina/emoglobina.xml"
 
-    EMBEDDINGS_PATH =  "./dataset/topo/embeddings"
-    GROUND_TRUE_PATH = "./dataset/topo/topo.xml"
+    # EMBEDDINGS_PATH =  "./dataset/topo/embeddings"
+    # GROUND_TRUE_PATH = "./dataset/topo/topo.xml"
     
     
     et = main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH)
@@ -368,7 +373,6 @@ if __name__ == "__main__":
     r = et.get_results()
 
[…]

### 2023-10-15 pipeline_phylogenetic.py

@@ -107,10 +107,10 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         function=pipeline_pca, # TODO t-sne
         list_args=[
             {"n_components": 10},
-            {"n_components": 20},
             {"n_components": 30},
-            {"n_components": 40},
             {"n_components": 50},
+            {"n_components": 70},
+            {"n_components": 90},
             {"n_components": "all"},       
         ]
     )
@@ -245,14 +245,14 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
 
 if __name__ == "__main__":
     
-    # EMBEDDINGS_PATH = "./dataset/emoglobina/embeddings"
-    # GROUND_TRUE_PATH = "./dataset/emoglobina/emoglobina.dist"
+    EMBEDDINGS_PATH = "./dataset/emoglobina/embeddings"
+    GROUND_TRUE_PATH = "./dataset/emoglobina/emoglobina.dist"
 
     # EMBEDDINGS_PATH =  "./dataset/batterio/embeddings"
     # GROUND_TRUE_PATH = "./dataset/batterio/batterio.dist"
 
-    EMBEDDINGS_PATH =  "./dataset/topo/embeddings"
-    GROUND_TRUE_PATH = "./dataset/topo/topo.dist"
+    # EMBEDDINGS_PATH =  "./dataset/topo/embeddings"
+    # GROUND_TRUE_PATH = "./dataset/topo/topo.dist"
     
     
     et = main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH)

### 2023-10-28 paper/paper.tex

@@
 As shown in Table \ref{tab:embedders}, the input length accepted is different between different models as well as the dimensionality of the output produced.
-The maximum length of the input sequence is usually determined by the length of the examples the model was training with or by limitations due to the structure of the model such as the size of the attention module in the models that use it.
+The maximum length of the input sequence is usually determined by the length of the examples the model was training on or by limitations due to the structure of the model such as the size of the attention module in the models that use it.
 These numerical representation of sequences are then used to various downstream applications.
@@
 Futhermore, if the sequence that we want to analyze is longer then the model's limit we have to split it in subsequences and thus the representation became a tensor.
+
+It is not clear which is the best way to aggregate this tensor in a fixed size representation.
 
@@
 
-\subsubsection{Similarity between distance matrices \label{sec:similarity}}
-In order to compare two distance matrices we performed an agglomerative clustering on both, the resulting tree is then cut at each level obtaining flat partitions of all possibles number of clusters.
-We performed a pairwise comparison of the partitions having the same number of clusters using the adjusted rand score \cite{hubert1985comparing}.
-The mean of these score, starting from two clusters up to $ \#elements - 1 $ clusters is called mean adjusted rand index (MARI).
-We compared different distance metrics as well as different methods to perform the hierarchical clustering.
-EXPLAIN BETTER
-
-\subsubsection{Enrichment analysis}
 The alignment distance matrix provide an evolutionary related distance between sequences \cite{SOFI202247}, we also wanted to analyze the properties of the embeddings at an higher level.
@@
 Both of them vary between $0$ and $1$, however $d1$ goes to 1 only when the two sets are equals while $d2$ goes to 1 also when one set is a subset of the other.
-After calculating one of these distance between all possible pair in the dataset we end up with a similarity matrix, that can be easily transformed in a distance matrix that is possible to compare with the distance matrix derived from the distance between the embeddings using the MARS as described in subsection \ref{sec:similarity}.
+After calculating one of these distance between all possible pair in the dataset we end up with a similarity matrix, that can be easily transformed in a distance matrix that is possible to compare with the distance matrix derived from the distance between the embeddings.
+
+\subsubsection{Similarity between distance matrices \label{sec:similarity}}
+In order to compare two distance matrices we performed an agglomerative clustering on both, the resulting tree is then cut at each level obtaining flat partitions of all possibles number of clusters.
+We performed a pairwise comparison of the partitions having the same number of clusters using the adjusted rand score \cite{hubert1985comparing}.
+The mean of these scores, starting from two clusters up to $ \#elements - 1 $ clusters is called mean adjusted rand index (MARI).
+We compared different distance metrics as well as different methods to perform the hierarchical clustering.
 

### 2023-12-18 pipeline_enrichment.py
bugfix (the scaling was performed only with pca != all)

@@ -74,7 +74,7 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
 
     # PRINCIPAL COMPONENT ANALYSIS (with scaling)
 
-    def pipeline_pca(previous_stage_output : dict, n_components) -> dict:
+    def pipeline_scaling_and_pca(previous_stage_output : dict, n_components) -> dict:
 
         """
         Performs PCA on the embeddings matrix after scaling it
@@ -90,9 +90,10 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         embeddings_matrix = previous_stage_output["embeddings_matrix"]
         embeddings_IDs = previous_stage_output["embeddings_IDs"]
 
+        scaler = StandardScaler()
+        embeddings_matrix = scaler.fit_transform(embeddings_matrix)
+
         if n_components != "all":
-            scaler = StandardScaler()
-            embeddings_matrix = scaler.fit_transform(embeddings_matrix)
             if n_components == "default":
                 n_components = min(embeddings_matrix.shape)
             pca = PCA(n_components=n_components)
@@ -101,13 +102,10 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         return { "embeddings_matrix" : embeddings_matrix, "embeddings_IDs": embeddings_IDs}
 
     et.add_multistage(
-        function=pipeline_pca,
+        function=pipeline_scaling_and_pca,
         list_args=[
             {"n_components": 10},
-            {"n_components": 30},
             {"n_components": 50},
-            {"n_components": 70},
-            {"n_components": 90},
             {"n_components": 'all'},
         ]
     )

### 2023-12-18 pipeline_phylogenetic.py

@@ -73,7 +73,7 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
 
     # PRINCIPAL COMPONENT ANALYSIS (with scaling)
 
-    def pipeline_pca(previous_stage_output : dict, n_components) -> dict:
+    def pipeline_scaling_and_pca(previous_stage_output : dict, n_components) -> dict:
 
         """
         Performs PCA on the embeddings matrix after scaling it
@@ -89,9 +89,10 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         embeddings_matrix = previous_stage_output["embeddings_matrix"]
         embeddings_IDs = previous_stage_output["embeddings_IDs"]
 
+        scaler = StandardScaler()
+        embeddings_matrix = scaler.fit_transform(embeddings_matrix)
+
         if n_components != "all":
-            scaler = StandardScaler()
-            embeddings_matrix = scaler.fit_transform(embeddings_matrix)
             if n_components == "default":
                 n_components = min(embeddings_matrix.shape)
             pca = PCA(n_components=n_components)
@@ -100,13 +101,10 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         return { "embeddings_matrix" : embeddings_matrix, "embeddings_IDs": embeddings_IDs}
 
     et.add_multistage(
-        function=pipeline_pca,
+        function=pipeline_scaling_and_pca,
         list_args=[
             {"n_components": 10},
-            {"n_components": 30},
             {"n_components": 50},
-            {"n_components": 70},
-            {"n_components": 90},
             {"n_components": "all"},       
         ]
     )

### 2023-12-18 autoembedding/results_manager.py
list of scores in the results

@@ -36,7 +36,7 @@ def results2table(
             if stage == "pipeline_build_embeddings_matrix":
                 embedder = args["embedder"]
                 combiner = args["combiner_method"]
-            if stage == "pipeline_pca":
+            if stage == "pipeline_scaling_and_pca":
                 pca = str(args["n_components"])
             if stage == "pipeline_build_embeddings_linkage_matrix":
                 metric_embedding = args["metric"]
@@ -95,6 +95,7 @@ def results2file(r, filepath):
     with open(filepath + ".txt", "w") as f:
         for result, pipeline in r:
             f.write(f"mean_adjusted_rand_score: {result['mean_adjusted_rand_score']} \n")
+            f.write(f"adjusted_rand_scores: {result['adjusted_rand_scores']} \n")
             for name, args in pipeline:
                 f.write(f"{name}  {args} \n")
             f.write(f"\n")

### 2023-12-18 pipeline_phylogenetic.py

@@ -47,10 +47,10 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
             {"embedder" : "seqvec", "combiner_method" : "sum" },
             {"embedder" : "seqvec", "combiner_method" : "max" },
             
-            {"embedder" : "dnabert", "combiner_method" : "pca" },
-            {"embedder" : "dnabert", "combiner_method" : "average" },
-            {"embedder" : "dnabert", "combiner_method" : "sum" },
-            {"embedder" : "dnabert", "combiner_method" : "max" },
+            # {"embedder" : "dnabert", "combiner_method" : "pca" },
+            # {"embedder" : "dnabert", "combiner_method" : "average" },
+            # {"embedder" : "dnabert", "combiner_method" : "sum" },
+            # {"embedder" : "dnabert", "combiner_method" : "max" },
             
             {"embedder" : "prose", "combiner_method" : "pca" },
             {"embedder" : "prose", "combiner_method" : "average" },
@@ -223,7 +223,7 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
         for i in range(predict_labels_matrix.shape[1]):
             adjusted_rand_scores.append(adjusted_rand_score(predict_labels_matrix[:,i], gtrue_labels_matrix[:,i]))
         
-        return {"mean_adjusted_rand_score" : np.mean(adjusted_rand_scores)}
+        return {"mean_adjusted_rand_score" : np.mean(adjusted_rand_scores), "adjusted_rand_scores": adjusted_rand_scores}
         
 
 

### 2023-12-19 pipeline_phylogenetic.py
bugfix

@@ -181,14 +181,14 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
 
     # COMPARE WITH GROUND TRUE
 
-    def pipeline_mean_adjusted_rand_score(previous_stage_output : dict, cluster_range : tuple ) -> dict:
+    def pipeline_mean_adjusted_rand_score(previous_stage_output : dict, cluster_range ) -> dict:
         """
         Computes the mean adjusted rand score between two hierarchical clustering averaging over all the cut
         in a given range
 
         Args:
             previous_stage_output (dict): The output of the previous stage, a dict containing the linkage matrix and the IDs
-            cluster_range (tuple): The range of clusters to consider (the cut to the hierarchical clustering)
+            cluster_range: The range of clusters to consider (the cut to the hierarchical clustering)
             ground_true_path (str): The path to the ground true newick file
         Returns:
             dict: A dict containing the mean adjusted rand score
@@ -200,19 +200,18 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
 
         if len(embeddings_IDs) != len(gtrue_IDs):
             raise Exception("The number of IDs in the ground true and the predicted clustering is different")
-        if cluster_range[0] < 2:
-            raise Exception("Cluster range must start at least from 2")
         if not set(embeddings_IDs) == set(gtrue_IDs):
             raise Exception("The IDs in the ground true and the predicted clustering are different")
 
-        start = cluster_range[0]
-        end = cluster_range[1]
-        
-        if end == -1:
-            end = min(len(embeddings_IDs), len(gtrue_IDs))
-        
-        predict_labels_matrix= cut_tree(embeddings_linkage_matrix, n_clusters=range(start, end))
-        gtrue_labels_matrix = cut_tree(gtrue_linkage_matrix, n_clusters=range(start, end))
+        start = 0
+        end = 0
+
+        if cluster_range == "auto":
+            start = 2
+            end = len(embeddings_IDs) - 1
+
+        predict_labels_matrix = cut_tree(embeddings_linkage_matrix, n_clusters=[i for i in range(start, end+1)])
+        gtrue_labels_matrix = cut_tree(gtrue_linkage_matrix, n_clusters=[i for i in range(start, end+1)])
 
         # order the matrix rows based on the IDs
[…]

### 2023-12-19 autoembedding/results_manager.py
list of ari

@@ -15,6 +15,8 @@ def results2table(
     ):
     
     computations_dict = {} # dict[combiner][pca][embedder] = score
+    best_result = -1
+    best_result_list = []
 
     # generate the empty dict with the argument that are relevant for the table
     for combiner in combiners:
@@ -69,6 +71,9 @@ def results2table(
             annotation == preferred_annotation:
         
             computations_dict[combiner][pca][embedder] = result[metric]
+            if result[metric] > best_result:
+                best_result = result[metric]
+                best_result_list = result['adjusted_rand_scores']
 
     
     data_matrix = [] # the matrix that will be converted to a dataframe, the order of the rows and columns must be consistent whith the one created by the MultiIndex and the "columns" parameter of the DataFrame constructor
@@ -86,7 +91,7 @@ def results2table(
     index = pd.MultiIndex.from_product(iterables, names=["combiner", "dimensional PCA"])
     df = pd.DataFrame( data_matrix, index=index, columns=embedders)
 
-    return df
+    return df, best_result_list
 
 def results2file(r, filepath):
     

### 2023-12-19 pipeline_enrichment.py
bugfix enri

@@ -4,7 +4,7 @@ from sklearn.decomposition import PCA
 from sklearn.preprocessing import StandardScaler
 from autoembedding.embeddings_matrix import build_embeddings_matrix
 from scipy.cluster.hierarchy import linkage
-from scipy.spatial.distance import pdist, squareform
+from scipy.spatial.distance import squareform
 from scipy.cluster.hierarchy import cut_tree
 from sklearn.metrics import adjusted_rand_score
 import numpy as np
@@ -48,10 +48,10 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
             {"embedder" : "seqvec", "combiner_method" : "sum" },
             {"embedder" : "seqvec", "combiner_method" : "max" },
             
-            {"embedder" : "dnabert", "combiner_method" : "pca" },
-            {"embedder" : "dnabert", "combiner_method" : "average" },
-            {"embedder" : "dnabert", "combiner_method" : "sum" },
-            {"embedder" : "dnabert", "combiner_method" : "max" },
+            # {"embedder" : "dnabert", "combiner_method" : "pca" },
+            # {"embedder" : "dnabert", "combiner_method" : "average" },
+            # {"embedder" : "dnabert", "combiner_method" : "sum" },
+            # {"embedder" : "dnabert", "combiner_method" : "max" },
             
             {"embedder" : "prose", "combiner_method" : "pca" },
             {"embedder" : "prose", "combiner_method" : "average" },
@@ -299,44 +299,54 @@ def main_et(EMBEDDINGS_PATH, GROUND_TRUE_PATH):
     )
 
 
-    def pipeline_mean_adjusted_rand_score(previous_stage_output : dict, cluster_range)-> dict:
+    def pipeline_mean_adjusted_rand_score(previous_stage_output : dict, cluster_range ) -> dict:
         """
-        Performs the enrichment analysis comparing the number of common annotations between sequences and the distance in the ebmeddings space
-        
-        Args:
-            previous_stage_output (dict): The output of the previous stage, a dict containing the embeddings matrix and the IDs
-            metric (str): The metric to use (euclidean, cosine, etc.)
-        Returns:
-            dict: A dict containing the linkage matrix and the IDs
-        """
+        Computes the mean adjusted rand score between two hierarchical clustering averaging over all the cut
+        in a given range
 
-        gtrue_linkage_matrix = previous_stage_output["gtrue_linkage_matrix"]
[…]

### 2023-12-19 autoembedding/ExecutionTree.py
upgrade in execution tree

@@ -21,6 +21,8 @@ class ExecutionTree:
         self.input = input
         self.root = Node(lambda previous_stage_output : previous_stage_output, {})
         self.frontier = [self.root]
+        self.num_nodes = 1
+        self.computed_nodes = 0
     
     def add_stage(self, function, args):
         
@@ -30,6 +32,7 @@ class ExecutionTree:
             new_n.parent = n
             n.childs.append(new_n)
             new_frontier.append(new_n)
+            self.num_nodes += 1
         
         self.frontier = new_frontier
 
@@ -43,6 +46,7 @@ class ExecutionTree:
                 new_n.parent = n
                 n.childs.append(new_n)
                 new_frontier.append(new_n)
+                self.num_nodes += 1
         
         self.frontier = new_frontier
     
@@ -50,6 +54,8 @@ class ExecutionTree:
     def recursive_dfs(self, node, previous_stage_output):
         
         current_output = node.compute(previous_stage_output=previous_stage_output)
+        self.computed_nodes += 1
+        print(f"Computed: {self.computed_nodes}/{self.num_nodes}")
         for c in node.childs:
             self.recursive_dfs(c, current_output)
 
